---
title: "R Notebook"
output:
  pdf_document: default
  html_notebook: default
  html_document:
    df_print: paged
---
Here we're try to find our is there a diffence in arr_delay among the different airports

Null Hypothesis (H0):
There is no significant difference in the mean "arr_delay" among the different airports.

Alternative Hypothesis (H1):
There is a significant difference in the mean "arr_delay" among at least some of the airports.

Explanation: 
H0: mu_1 = mu_2 = mu_3 = ... = mu_k (where mu represents the mean "arr_delay" for each airport)
H1: At least one pair of means mu_i and mu_j is different.
Here, mu_1, mu_2, ..., mu_k represent the mean "arr_delay" for each airport category. If the p-value is less than the significance level (commonly used in our Stat501 class:  0.05), we reject the null hypothesis and conclude that there is evidence to suggest that at least some of the means are different.

Loading the data set only for 4 Top Airports: AUS, MSP, BNA, PIT
```{r}
library(readr)
raw_top4_airport <- read_csv("~/Documents/Stat501/project/raw_top4_airport.csv")
```


```{r}
library(ggplot2)

# Create a box plot
ggplot(raw_top4_airport, aes(x = airport, y = arr_delay)) +
  geom_boxplot(fill = "skyblue", color = "black") +
  labs(title = "Box Plot of arr_delay by Carrier",
       x = "Carrier",
       y = "arr_delay") +
  theme_minimal()
```
Through this box plot, we can see that we have some outliers. So, the we're making ANOVA analysis with and without Outlier. 

a) With Outliers, the Anova results are:
```{r}

library(tidyverse)

anova_result_w_outliers <- aov(arr_delay ~ airport, data = raw_top4_airport)

print(summary(anova_result_w_outliers))

```
Interpretation:

Df (Degrees of Freedom): There are three degrees of freedom for the factor "airport" and 741 degrees of freedom for residuals.
Sum Sq (Sum of Squares): This represents the sum of squared differences between the observed values and the mean. For "airport," it is 2.789e+09, and for residuals, it is 9.642e+10.
Mean Sq (Mean Square): Mean Squares are calculated by dividing the Sum of Squares by the corresponding degrees of freedom.
F value: The F statistic is a ratio of the variance between groups to the variance within groups. Here, it is 7.144.
Pr(>F): This is the p-value associated with the F statistic. It is extremely small (9.84e-05), indicating that there is a significant difference in mean "arr_delay" among at least two airports.

Since we are rejecting the Null Hypothesis, we want to investigae fruther. The method used for investigating the pairs is Tukey's post-hoc test:

```{r}
# Perform Tukey's post-hoc test
tukey_result_w_outliers <- TukeyHSD(anova_result_w_outliers)

# Display Tukey's post-hoc results
print(tukey_result_w_outliers)
```

diff (Difference): The estimated difference in means between the pairs of airports for "arr_delay."
lwr and upr (Lower and Upper Confidence Intervals): The 95% confidence interval for the difference in means.
p adj (Adjusted p-value): The p-value adjusted for multiple comparisons (Tukey's correction).
Interpretation:
BNA-AUS: The difference in mean "arr_delay" between Nashville (BNA) and Austin (AUS) is not statistically significant (p = 0.9918).

MSP-AUS: The difference in mean "arr_delay" between Minneapolis (MSP) and Austin (AUS) is not statistically significant (p = 0.8366).

PIT-AUS: The difference in mean "arr_delay" between Pittsburgh (PIT) and Austin (AUS) is statistically significant (p = 0.00299). The negative difference suggests that Austin has a higher mean delay than Pittsburgh.

MSP-BNA: The difference in mean "arr_delay" between Minneapolis (MSP) and Nashville (BNA) is not statistically significant (p = 0.6758).

PIT-BNA: The difference in mean "arr_delay" between Pittsburgh (PIT) and Nashville (BNA) is statistically significant (p = 0.00812). The negative difference suggests that Nashville has a higher mean delay than Pittsburgh.

PIT-MSP: The difference in mean "arr_delay" between Pittsburgh (PIT) and Minneapolis (MSP) is statistically significant (p = 0.0001087). The negative difference suggests that Minneapolis has a higher mean delay than Pittsburgh.

In summary, the adjusted p-values indicate whether the differences in mean "arr_delay" are statistically significant after correcting for multiple comparisons.


b) We are eager to know what effect if we remove outliers
So, after removing outlier, we performed same analysis again.
```{r}
# Load the required libraries
library(ggplot2)

# Function to remove outliers
remove_outliers <- function(x) {
  q <- quantile(x, c(0.25, 0.75))
  iqr <- q[2] - q[1]
  lower_bound <- q[1] - 1.5 * iqr
  upper_bound <- q[2] + 1.5 * iqr
  return(x[x >= lower_bound & x <= upper_bound])
}

# Remove outliers from 'arr_delay'
flight_data_no_outliers <- raw_top4_airport %>%
  filter(!arr_delay %in% remove_outliers(raw_top4_airport$arr_delay))

# Create a box plot without outliers
ggplot(flight_data_no_outliers, aes(x = airport, y = arr_delay)) +
  geom_boxplot(fill = "skyblue", color = "black") +
  labs(title = "Box Plot of arr_delay by Carrier (Without Outliers)",
       x = "Carrier",
       y = "arr_delay") +
  theme_minimal()
```

```{r}
# Load the required libraries
library(tidyverse)

# Assuming your data is stored in a dataframe named flight_data_no_outliers
# Perform one-way ANOVA
anova_result_no_outliers <- aov(arr_delay ~ airport, data = flight_data_no_outliers)

# Display ANOVA results
print(summary(anova_result_no_outliers))

```

The p-value (0.00584) is less than 0.05, indicating that there is a significant difference in mean "arr_delay" among at least two airports.

Tukey Multiple Comparisons:

```{r}
# Perform Tukey's post-hoc test
tukey_result <- TukeyHSD(anova_result_no_outliers)

# Display Tukey's post-hoc results
print(tukey_result)
```

The p-value for the pair MSP-AUS is 0.0146183 and PIT-MSP, that are less than 0.05. This suggests a significant difference in mean "arr_delay" between Minneapolis (MSP) and Austin (AUS), PIT and MSP respectively .



